Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-10-18 inference.tex
Encoding algorithm.

non-negatives ($I(\stg) \Leftrightarrow \forall(r\in\stg).~r.\C{bal}\ge

@@

account database that is equal to $a$
($\exists(r\in\stg).~r=a$). However, this assertion is unstable
because a concurrent \C{withdraw} or \C{deposit} transactions might
update the account balance, so that such a record no longer exists in
the database.  Fortunately, the weakening operator
($\stabilize{\cdot}$) allows us to weaken the assertion to $\exists
\stg'.I(\stg') \conj \exists(r\in\stg').~a=r$ (Fig.~\ref{fig:logic}
formalizes the encoding of {\sf exists} to logic), which is enough to
prove that $\C{a.bal + pc*a.bal}\ge 0$, and verify that
\C{add\_interest} preserves the positive balance invariant.

@@

The correspondence between the inference rules and the RG judgment of
\S~\ref{sec:reasoning} is stated thus\footnote{The proof can be found in
the supplementary material.}:
\begin{theorem}
\label{thm:inference-sound}
  \emph{Forall} i,$R$,$I$,$c$,$\F$, if $\stable(\R,I)$ and $c \elabsto \F$,
  then:\\\vspace*{-0.2cm}
  \begin{smathpar}
  \begin{array}{c}
  \R \vdash \hoare{\lambda(\stl,\stg).~\stl=\emptyset \conj
  I(\stg)}{c}{\lambda(\stl,\stg).\stl \subseteq \F(\emptyset,\stg)}
  \end{array}
  \end{smathpar}
\end{theorem}
\noindent where the set expression $\F(\emptyset,\stg)$ has a natural
interpretation as discussed below.
% Where, ${\sf true} = \lambda(\stl,\stg).~true$ and $\stable(\R,I)
% \Leftrightarrow \forall\stl,\stg,\stg'. I(\stg) \conj
% \R(\stl,\stg,\stg') \Rightarrow \I(\stg')$.

@@

Theorem~\ref{thm:inference-sound} lets us replace the local judgment
of the \rulelabel{RG-Txn} rule (Fig.~\ref{fig:rg-rules}) by a state
transformer inference judgment. The soundness of a transaction's guarantee can now
be established w.r.t the effect $\F$ of the body. The
\rulelabel{RG-Txn} rule so updated is shown
below:
\begin{smathpar}
\begin{array}{c}
\RULE
{
  \stable(R,\I)\spc
  \stable(R,I)\\
   \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \spc  c \elabsto \F\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg') \spc \stable(\R_c,\F)\\
  \forall \stg.~ G(\stg, F(\emptyset,\stg) \gg \stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G,I}
}
\end{array}
\end{smathpar}
Automating the application of the \rulelabel{RG-Txn} rule for a
transaction requires automating the multiple implication checks in
the premise. While $R$, $\R$, $\R_c$, $\I$ and $I$ are formulas in
first-order logic (FOL) with a relatively simple structure, $\F$
is an expression in the set language $\SL$
(Fig.~\ref{fig:logic-syntax}) with a possibly complex structure.
Fortunately, however, there exists a semantics-preserving translation
from $\SL$ to a restricted subset of first-order logic (FOL) that
lends itself to automatic reasoning.

>>>>

non-negative ($I(\stg) \Leftrightarrow \forall(r\in\stg).~r.\C{bal}\ge

@@

account database that is equal to $a$ ($\exists(r\in\stg).~r=a$).
However, this assertion is unstable because a concurrent \C{withdraw}
or \C{deposit} transactions might update the account balance, so that
such a record no longer exists in the database.  Fortunately, the
weakening operator ($\stabilize{\cdot}$) allows us to weaken the
assertion to $\exists \stg'.I(\stg') \conj \exists(r\in\stg').~a=r$
({\sf exists} translates to $\exists$ in first-order logic, as
formalized in Fig.~\ref{fig:logic}), which is enough to prove that
$\C{a.bal + pc*a.bal}\ge 0$, and verify that \C{add\_interest}
preserves the positive balance invariant.

@@

The correspondence between the inference rules and the RG judgment of
\S~\ref{sec:reasoning} is stated thus\footnote{The proof can be found in
the supplementary material.}:
\begin{theorem}
\label{thm:inference-sound}
  \emph{Forall} i,$R$,$I$,$c$,$\F$, if $\stable(\R,I)$ and $c \elabsto \F$,
  then:\\\vspace*{-0.2cm}
  \begin{smathpar}
  \begin{array}{c}
  \R \vdash \hoare{\lambda(\stl,\stg).~\stl=\emptyset \conj
  I(\stg)}{c}{\lambda(\stl,\stg).\stl = \F(\emptyset,\stg)}
  \end{array}
  \end{smathpar}
\end{theorem}
\noindent where the set expression $\F(\emptyset,\stg)$ has a natural
interpretation in the theory of sets.
% Where, ${\sf true} = \lambda(\stl,\stg).~true$ and $\stable(\R,I)
% \Leftrightarrow \forall\stl,\stg,\stg'. I(\stg) \conj
% \R(\stl,\stg,\stg') \Rightarrow \I(\stg')$.

@@

Theorem~\ref{thm:inference-sound} lets us replace the local judgment
of the \rulelabel{RG-Txn} rule (Fig.~\ref{fig:rg-rules}) by a state
transformer inference judgment. The soundness of a transaction's
guarantee can now be established w.r.t the effect $\F$ of the body.
The \rulelabel{RG-Txn} rule so updated is shown below:
\begin{smathpar}
\begin{array}{c}
\RULE
{
  \stable(R,\I)\spc
  \stable(R,I)\spc
  \R_e(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_e(\stl, \stg_1, \stg') \spc  c \Longrightarrow_{\langle i,\R_e,I \rangle}\F\\
  \R_c(\stl,\stg,\stg') \Leftrightarrow \exists \stg_1.  R(\stg, \stg') \wedge \I_c(\stl, \stg_1, \stg') \spc \stable(\R_c,\F)\\
  \forall \stl,\stg.~ \stl \in F(\emptyset,\stg) \Rightarrow 
    G(\stg, \stl\gg\stg)\spc
  \forall \stg,\stg'.~I(\stg) \wedge G(\stg,\stg') \Rightarrow I(\stg')\\
}
{
  \rg{I,R}{\ctxn{i}{\I}{c}}{G,I}
}
\end{array}
\end{smathpar}
Automating the application of the \rulelabel{RG-Txn} rule for a
transaction requires automating the multiple implication checks in
the premise. While $R$, $\R$, $\R_c$, $\I$ and $I$ are formulas in
first-order logic (FOL) with a relatively simple structure, $\F$
is an expression in the set language $\SL$
(Fig.~\ref{fig:logic-syntax}) with a possibly complex structure.
Fortunately, however, there exists a semantics-preserving translation
from $\SL$ to a restricted subset of first-order logic (FOL) that
lends itself to automatic reasoning.
